An Experimental Evaluation of Hybrid Query Algorithms

### 2025-04-01 VLDB2026/main.tex
Update main.tex

\documentclass[sigconf, nonacm]{acmart}
\usepackage{fontspec}
\usepackage{balance}
\usepackage{xeCJK}
\usepackage{enumitem}
\usepackage{multirow}
\usepackage{svg} % For SVG support
\usepackage{graphicx}
\usepackage{microtype}  %处理换行

@@

Nearest Neighbor (NN) search \cite{cover1967nearest}aims to find the closest vector in a given space and serves as a fundamental algorithm in vector retrieval, with widespread applications in recommendation systems  \cite{monolith, Airbnb, Related_Pins_at_Pinterest, wang2018billion, pal2020pinnersage, okura2017embedding}  and image retrieval \cite{wang2012scalable}.
% 需要添加引用，下面这段
However, with the exponential growth of data volume and the increasing dimensionality of vectors \cite{weber1998quantitative}, traditional NN search methods struggle to meet the demands of real-time search. To address this issue, researchers have turned to more efficient approaches—Approximate Nearest Neighbor search ~\cite{arya1998optimal,beis1997shape,gionis1999similarity,malkov2018efficient}. ANN algorithms significantly improve search efficiency by constructing effective indexing structures, albeit at the cost of reduced accuracy.

@@

（1）\textbf{ Systematic Classification and Overview.}
We systematically classify more than ten representative attribute filtering algorithms from multiple dimensions, including index structure, filtering strategy, Boolean logic support, and index construction methods. Additionally, we survey range filtering algorithms and provide an overview of widely used vector retrieval libraries (e.g., Faiss) and vector databases (e.g., Milvus, PASE, VBASE). These efforts provide a unified reference framework for future research.

（2）\textbf{ Evaluation of Attribute-Filtering Algorithms.}
% We select several representative algorithms for Attribute Filtered ANN Search. Using six real-world datasets, we evaluate their performance under different settings, such as the number of attributes, attribute distribution, and selectivity. We analyze their performance in terms of index construction time, index size,peak memory usage, query speed (QPS), and query accuracy. We also discuss their robustness and adaptability in complex scenarios.
We experimentally evaluate 13 attribute filtering ANN algorithms on six real-world datasets. By analyzing the performance variations of these algorithms under different numbers of attributes, attribute distributions, and selectivity conditions, we reveal the strengths and weaknesses of each algorithm. Furthermore, we explore the robustness and adaptability of these algorithms under complex query conditions. Evaluation metrics include index construction time, index size, peak memory usage,QPS and search accuracy.

（3）\textbf{ Evaluation of Range-Filtering Algorithms.}
% We evaluate five main algorithms (SeRF, WinFilter, iRangeGraph, DSG, and UNIFY) for Range Filtered ANN Search. We run experiments on three large-scale datasets with different query range settings. We compare their performance in index construction, storage usage, and query efficiency. We also study the effect of index structures on algorithm performance. This provides useful insights for future improvements.
We benchmark five mainstream range filtering ANN algorithms on three large-scale datasets, using varying query range settings in the experiments. The results demonstrate the performance of these algorithms in terms of index construction efficiency, storage overhead, and query performance. Additionally, we conduct an in-depth analysis of how index structures impact algorithm performance, providing valuable insights for algorithm design.

（4）\textbf{ Recommendations and Challenges.}

Based on the experimental results, we provide algorithm selection recommendations for common application scenarios and highlight key challenges in the field of hybrid queries. These challenges include limited Boolean logic support, the lack of multi-attribute range filtering capabilities, the high indexing costs of graph-based methods, and the sensitivity of algorithms to data distribution. Currently, few methods simultaneously support both attribute filtering and range filtering, pointing to potential directions for future research.

@@

\vspace{1em}
\noindent\textbf{\underline{UNG \cite{UNG}.}} 
Graph-based methods（NHQ, Filtered-DiskANN, ACO-\\RN）do not guarantee result completeness, while CAPS lacks robustness when base and query vectors differ in attribute cardinality. UNG is proposed to overcome both limitations.

@@

\vspace{1em}
\noindent\textbf{\underline{DSG \cite{DSG}.}}  
Most RF-ANN search methods, such as iRange and WinFilter, are designed for static datasets. SeRF allows incremental insertion but requires attributes to be ordered. DSG introduces the first truly dynamic RF-ANN framework that supports inserting data with unordered attributes while maintaining efficient range filtering.
DSG is built on two key data structures. 1) The rectangle tree partitions the query space into rectangular regions, each corresponding to a group of queries that share the same nearest neighbors. 2) The dynamic segment graph is a neighbor graph where each edge is annotated with the attribute range in which it is valid.
With these structures, only a few regions need to be updated when inserting new data. During queries, the system considers only edges valid for the current attribute range, improving efficiency.

@@

    \begin{subfigure}{\columnwidth}
        \centering
        \setlength{\abovecaptionskip}{0cm}
        \setlength{\belowcaptionskip}{-0.3cm}
        \includegraphics[width=\linewidth]{figures/indexData/exp_7_build_time_comparison_query1.pdf}
        \caption{Index construction time}
        \label{fig:build_time_comparison_query1}
    \end{subfigure}

@@

\vspace{0.5em}

@@

\vspace{0.5em}
\textit{\textbf{Index construction time.}}
As shown in Figure~\ref{fig:build_time_comparison_query1}, among the aforementioned four datasets, PASE has the longest index construction time due to its single-threaded indexing support. Additionally, VBASE also supports only single-threaded indexing but achieves better query performance than PASE.

@@

\vspace{0.5em}
\textit{\textbf{Index size.}}
Traditional graph-based methods often only evaluate the size of the generated graph index and do not include the original data, while the index generated by the IVF-based method includes the original data. In order to unify the comparison standard, we include both the index structure generated by the algorithm and the original data when evaluating the index size. As shown in the figure~\ref{fig:index_size_mb_comparison_query1}, PASE exhibits significantly larger index sizes across all supported datasets. This overhead stems primarily from its extensive metadata requirements. Other methods demonstrate comparable index sizes, with Faiss consistently achieving the most compact representations through its efficient index design.

\vspace{0.5em}
\textit{\textbf{Peak memory usage.}}
As shown in Figure~\ref{fig:memory_mb_comparison_query1}, on small-scale datasets, all methods except NHQ exhibit low memory usage. The substantially higher memory consumption of NHQ is attributed to its graph construction procedure, which requires loading the entire dataset into memory. On large-scale datasets, CPAS exhibits slightly lower memory usage compared to other algorithms. Due to their special characteristics, databases were not included in the peak memory usage testing during index construction.

@@

\begin{figure}
    \centering
    \setlength{\abovecaptionskip}{0cm}
    \setlength{\belowcaptionskip}{-0.1cm}
    % 上面的图例，居中并可通过 hspace 调整左右位置
    \hspace*{15pt} % 可调整的参数，负值左移，正值右移，例如 -20pt 或 20pt
    \includegraphics[width=0.98\columnwidth]{figures/exp/exp_2_legend.pdf} % 图例图片路径
    \vspace{-10pt} % 使用负值拉近两张图的距离，可以根据需要调整

    % 下面的子图，居中显示
    \includegraphics[width=0.9\columnwidth]{figures/exp/exp_2_1.pdf}
    \caption{Effect of Multi-Attribute Index Construction on Single Query Performance. ($S$ denotes single-attribute index construction and single-attribute query; $M$ denotes multi-attribute index construction and single-attribute query)}
    \label{fig:exp_2_1}

@@

\vspace{1em}
\textit{\textbf{Multi-Attribute Building and Single-Attribute Query.}}
In this scenario, the index is constructed using multiple attributes but applies filtering condition on only one attribute during query processing.

@@

@@

\vspace{10pt}
\textit{\textbf{Multi-Attribute Building and Multi-Attribute Query.}}  
Compared to single-attribute filtering, multi-attribute joint filtering better reflects real-world application scenarios. To evaluate algorithm performance under such scenario, we construct the index using three uniformly distributed attributes and apply filtering conditions on all three attributes simultaneously during the query process.

@@

The IVF-based algorithm CAPS, optimized for hybrid query scenarios, employs a multi-level partitioning strategy that introduces additional overhead. This overhead negatively impacts performance on small datasets. However, on large datasets, the overhead is amortized, and the benefits of attribute-aware partitioning become evident, making CAPS the second-best performer after UNG.

@@

%4.2.7
\textit{\textbf{Different Datasets .}}  
To evaluate the performance of various methods across different datasets, we conduct experiments on six datasets. Our analysis focuses on performance variations concerning dataset size, LID, and vector dimensionality.

@@

\begin{table}[htbp]
\centering
\setlength{\abovecaptionskip}{0.05cm}
\caption{Time and space overhead of range filtering index construction}
% \vspace{-7pt}
\label{tab:Range Filtered}
\resizebox{\columnwidth}{!}{
\begin{tabular}{l l c c c c c }
\toprule
\textbf{Dataset} & \textbf{Metric} & \textbf{DSG} & \textbf{IRange} & \textbf{SeRF} & \textbf{UNIFY} & \textbf{WinFilter} \\
\midrule
\multirow{3}{*}{DEEP} 
& Index time(s)  & 4156 & 1128 & \textbf{189.89} & 7510 & 32843 \\
& Index size(GB)  &  4.6 & 0.92 & \textbf{0.80} & 1.43 & 2.41  \\
& Peak memory(GB) &  7.78 & 3.54 & \textbf{1.38} & 2.61 & 7.08   \\
% \cmidrule{1-7}
\midrule
\multirow{3}{*}{YT-Audio} 
& Index time(s)   & 15314 & 1039 & \textbf{187.20} & 5905 & 30904    \\
& Index size(GB)    & 4.3 & \textbf{0.7} & 1.04 & 1.55 & 1.32  \\
& Peak memory(GB)   & 7.5 & 3.16 & \textbf{1.57} & 2.97 & 6.95  \\
\midrule
\multirow{3}{*}{WIT} 
& Index time(s)  & 29492 & 7068 & \textbf{828.12} & 28759 & 94909\\
& Index size(GB) &  4.2 & \textbf{1.5 }& 15.34 & 8.71 & 1.59  \\
& Peak memory(GB)  & 21.7 & \textbf{12.32} & 15.87 & 24.38 & 15.29\\
% \cmidrule{1-7}

\bottomrule
\end{tabular}
}
\end{table}
Similar to the evaluation of \textit{Attribute Filtering} algorithms, we also evaluate three key metrics mentioned above in \textit{Range Filtering} algorithms.

\vspace{0.5em}
\textit{\textbf{Index construction time.}}
As shown in Table \ref{tab:Range Filtered}, SeRF achieves the fastest index construction time among the five algorithms. It incrementally builds the index in attribute order. It requires only a single pass and adds edges within local graphs. The structure is simple, with no redundant operations.

@@

\vspace{0.5em}

\textit{\textbf{Index size.}} Across the three datasets, IRange maintains a relatively small index size. It builds independent local neighbor graphs within each interval and merges indices dynamically during queries. This prevents space explosion. On low-dimensional datasets (Deep, YT-Audio), SeRF compresses the index using sparse vector distributions, keeping it small. However, in the high-dimensional WIT dataset with high LID, vector connectivity increases. This reduces compression efficiency, leading to a sharp rise in index size. DSG maintains a stable index size across datasets but is generally larger. It uses edge structures with interval labels to support streaming insertions and range queries, increasing storage overhead.
\vspace{0.5em}

\textit{\textbf{Peak memory usage.}}

In terms of peak memory usage, all five algorithms consume less memory on the first two datasets. However, memory usage increases significantly on WIT, likely due to its high dimensionality.

@@

In contrast, IRange achieves the lowest peak memory usage on WIT. It pre-builds graphs only for a limited number of intervals, avoiding explicit index construction for all possible query ranges. Its index structure remains stable in high dimensions without significant graph growth. This effectively controls index size, making IRange more advantageous for high-dimensional datasets.
\vspace{0.5em}

\begin{figure*}[htbp]
  \centering
  \setlength{\abovecaptionskip}{0cm}
    \setlength{\belowcaptionskip}{-0.4cm}
  \includegraphics[width=\textwidth]{figures/exp/exp_8_2.pdf}
  \caption{RF-ANN search }
  \label{fig:exp_8_2}
\end{figure*}
\subsubsection{Performance Evaluation. }

@@

\begin{figure*}[htbp]
    \centering
    \setlength{\abovecaptionskip}{0cm}
    \setlength{\belowcaptionskip}{-0.4cm}
    \includegraphics[width=\textwidth]{figures/exp/exp_8_3.pdf}
    \caption{Dataset Effect on RF-ANN Search}
    \label{fig:exp_8_3}
\end{figure*}

@@

\begin{table}[htbp]
\centering
\setlength{\abovecaptionskip}{0.05cm}
\caption{Algorithm Recommendation per Scenario}
\resizebox{\columnwidth}{!}{
\begin{tabular}{|l|c|c|}
\hline
\textbf{Scenario} & \textbf{Attribute Filtering} & \textbf{Range Filtering}\\
\hline
S1: Large-scale Datasets & Puck, UNG & iRange\\
\hline
S2: Fast Index Construction & ACORN-1, UNG  &SeRF, iRange\\
\hline
S3: Query on Hard Datasets & CAPS, UNG &iRange, DSG\\
\hline
S4: Query on Easy Datasets & StitchedVamana, NHQ & iRange\\

\hline
S5: General-purpose Use & Milvus, Faiss &DSG, UNIFY\\
\hline
S6: Resource-constrained Environments & CAPS, Faiss&SeRF, IRange \\
\hline
S7: High-Recall Querying & UNG & WinFilter \\
\hline
\end{tabular}
}
\label{tab:algo_scenarios}
\vspace{-0.3cm}
\end{table}

@@

% \end{tabular}
% \label{tab:range_algo_scenarios}
% \end{table}
Table~\ref{tab:algo_scenarios} also presents the application scenarios for range filtering. iRange demonstrates strong performance on large-scale datasets, combining efficient construction with excellent query capabilities (S1). For fast index construction, SeRF leads with iRange as a close second (S2). Both SeRF and iRange show superior memory efficiency (S6). When handling hard datasets, iRange and DSG are recommended (S3), while iRange also performs well on easy datasets (S4). For general-purpose applications, DSG supports dynamic updates while UNIFY provides flexible filtering strategies (S5). WinFilter emerges as the optimal choice for high-recall static queries (S7).

\subsection{Challenges}
\subsubsection{\textbf{Attribute Filtering}}
%修改语序
Most existing attribute filtering algorithms are graph-based, and they typically outperform other methods in terms of query accuracy and efficiency. However, graph-based algorithms often incur high indexing costs and significant memory overhead. In contrast, IVF algorithms  generally consume less memory but suffer from limited performance. Filtered−DiskANN leverages disk-based storage, making it more suitable for extremely large-scale datasets even in memory-constrained environments. Moreover, the index construction phase of graph-based algorithms is computationally intensive. Therefore, exploring GPU-based acceleration for graph construction and vector computation during indexing is promising. It can significantly improve the performance of graph-based methods.

>>>>

\documentclass[sigconf, nonacm]{acmart}
\usepackage{fontspec}
\usepackage{balance}
% \usepackage{xeCJK}
\usepackage{enumitem}
\usepackage{multirow}
% \usepackage{svg} % For SVG support
\usepackage{graphicx}
\usepackage{microtype}  %处理换行

@@

Nearest Neighbor (NN) search \cite{cover1967nearest} aims to find the closest vector in a given space and serves as a fundamental algorithm in vector retrieval, with widespread applications in recommendation systems  \cite{monolith, Airbnb, Related_Pins_at_Pinterest, wang2018billion, pal2020pinnersage, okura2017embedding}  and image retrieval \cite{wang2012scalable}.
% 需要添加引用，下面这段
However, with the exponential growth of data volume and the increasing dimensionality of vectors \cite{weber1998quantitative}, traditional NN search methods struggle to meet the demands of real-time search. To address this issue, researchers have turned to more efficient approaches—Approximate Nearest Neighbor search ~\cite{arya1998optimal,beis1997shape,gionis1999similarity,malkov2018efficient}. ANN algorithms significantly improve search efficiency by constructing effective indexing structures, albeit at the cost of reduced accuracy.

@@

(1)\textbf{ Systematic Classification and Overview.}
We systematically classify more than ten representative attribute filtering algorithms from multiple dimensions, including index structure, filtering strategy, Boolean logic support, and index construction methods. Additionally, we survey range filtering algorithms and provide an overview of widely used vector retrieval libraries (e.g., Faiss) and vector databases (e.g., Milvus, PASE, VBASE). These efforts provide a unified reference framework for future research.

(2)\textbf{ Evaluation of Attribute-Filtering Algorithms.}
% We select several representative algorithms for Attribute Filtered ANN Search. Using six real-world datasets, we evaluate their performance under different settings, such as the number of attributes, attribute distribution, and selectivity. We analyze their performance in terms of index construction time, index size,peak memory usage, query speed (QPS), and query accuracy. We also discuss their robustness and adaptability in complex scenarios.
We experimentally evaluate 13 attribute filtering ANN algorithms on six real-world datasets. By analyzing the performance variations of these algorithms under different numbers of attributes, attribute distributions, and selectivity conditions, we reveal the strengths and weaknesses of each algorithm. Furthermore, we explore the robustness and adaptability of these algorithms under complex query conditions. Evaluation metrics include index construction time, index size, peak memory usage,QPS and search accuracy.

(3)\textbf{ Evaluation of Range-Filtering Algorithms.}
% We evaluate five main algorithms (SeRF, WinFilter, iRangeGraph, DSG, and UNIFY) for Range Filtered ANN Search. We run experiments on three large-scale datasets with different query range settings. We compare their performance in index construction, storage usage, and query efficiency. We also study the effect of index structures on algorithm performance. This provides useful insights for future improvements.
We benchmark five mainstream range filtering ANN algorithms on three large-scale datasets, using varying query range settings in the experiments. The results demonstrate the performance of these algorithms in terms of index construction efficiency, storage overhead, and query performance. Additionally, we conduct an in-depth analysis of how index structures impact algorithm performance, providing valuable insights for algorithm design.

(4)\textbf{ Recommendations and Challenges.}
Based on the experimental results, we provide algorithm selection recommendations for common application scenarios and highlight key challenges in the field of hybrid queries. These challenges include limited Boolean logic support, the lack of multi-attribute range filtering capabilities, the high indexing costs of graph-based methods, and the sensitivity of algorithms to data distribution. Currently, few methods simultaneously support both attribute filtering and range filtering, pointing to potential directions for future research.

@@

\vspace{1em}
\noindent\textbf{\underline{UNG \cite{UNG}.}} 
Graph-based methods(NHQ, Filtered-DiskANN, ACO-\\RN)do not guarantee result completeness, while CAPS lacks robustness when base and query vectors differ in attribute cardinality. UNG is proposed to overcome both limitations.

@@

\vspace{1em}
\noindent\textbf{\underline{DSG \cite{DSG}.}}  
Most RF-ANN search methods, such as iRange and WinFilter, are designed for static datasets. SeRF allows incremental insertion but requires ordered attributes. DSG introduces the first dynamic RF-ANN framework supporting data insertion with unordered attributes while maintaining efficient range filtering.
DSG relies on two key data structures: 1) A rectangle tree partitions query space into rectangular regions, each corresponding to a group of queries sharing nearest neighbors. 2) A dynamic segment graph is a neighbor graph where each edge is annotated with its valid attribute range.
With these structures, only few regions need updates when inserting new data. During queries, the system considers only edges valid for current attribute range, boosting efficiency.

@@

    \begin{subfigure}{\columnwidth}
        \centering
        \setlength{\abovecaptionskip}{0cm}

        \setlength{\belowcaptionskip}{-0.3cm}
        \includegraphics[width=\linewidth]{figures/indexData/exp_7_build_time_comparison_query1.pdf}
        \caption{Index construction time}
        \label{fig:build_time_comparison_query1}
    \end{subfigure}

@@

@@

\textit{\textbf{Index construction time.}}
As shown in Figure~\ref{fig:build_time_comparison_query1}, among the aforementioned four datasets, PASE has the longest index construction time due to its single-threaded indexing support. Additionally, VBASE also supports only single-threaded indexing but achieves better query performance than PASE.

@@

\textit{\textbf{Index size.}}
Traditional graph-based methods often only evaluate the size of the generated graph index and do not include the original data, while the index generated by the IVF-based method includes the original data. In order to unify the comparison standard, 
we include both the index structure generated by the algorithm and the original data when evaluating the index size. 
% Traditional graph-based methods often evaluate only the size of the generated index, excluding the original data, whereas IVF-based methods include it. To ensure a consistent comparison, we consider both the index structure and the original data when measuring index size.
As shown in the figure~\ref{fig:index_size_mb_comparison_query1}, PASE exhibits significantly larger index sizes across all supported datasets. This overhead stems primarily from its extensive metadata requirements. The index sizes of other methods show minimal variation. Among these, Faiss generates slightly smaller indexes due to its efficient index design.

\textit{\textbf{Peak memory usage.}}
As shown in Figure~\ref{fig:memory_mb_comparison_query1}, on small-scale datasets, all methods except NHQ exhibit low memory usage. The substantially higher memory consumption of NHQ is attributed to its graph construction procedure, which requires loading the entire dataset into memory. On large-scale datasets, CPAS exhibits slightly lower memory usage compared to other algorithms. Due to their special characteristics, databases were not included in the peak memory usage testing during index construction.

@@

@@

\textit{\textbf{Multi-Attribute Building and Single-Attribute Query.}}
In this scenario, the index is constructed using multiple attributes but applies filtering condition on only one attribute during query processing.

@@

\begin{figure}[th]
    \centering
    \setlength{\abovecaptionskip}{0cm}
    \setlength{\belowcaptionskip}{-0.1cm}
    % 上面的图例，居中并可通过 hspace 调整左右位置
    \hspace*{15pt} % 可调整的参数，负值左移，正值右移，例如 -20pt 或 20pt
    \includegraphics[width=0.98\columnwidth]{figures/exp/exp_2_legend.pdf} % 图例图片路径
    \vspace{-10pt} % 使用负值拉近两张图的距离，可以根据需要调整

    % 下面的子图，居中显示
    \includegraphics[width=0.8\columnwidth]{figures/exp/exp_2_1.pdf}
    \caption{Effect of Multi-Attribute Index Construction on Single Query Performance. ($S$ denotes single-attribute index construction and single-attribute query; $M$ denotes multi-attribute index construction and single-attribute query)}
    \label{fig:exp_2_1}

@@

\textit{\textbf{Multi-Attribute Building and Multi-Attribute Query.}}  
Compared to single-attribute filtering, multi-attribute joint filtering better reflects real-world application scenarios. To evaluate algorithm performance under such scenario, we construct the index using three uniformly distributed attributes and apply filtering conditions on all three attributes simultaneously during the query process.

@@

The IVF-based algorithm CAPS, optimized for hybrid query scenarios, employs a multi-level partitioning strategy that introduces additional overhead. This overhead negatively impacts performan-\\ce on small datasets. However, on large datasets, the overhead is amortized, and the benefits of attribute-aware partitioning become evident, making CAPS the second-best performer after UNG.

@@

%4.2.7

\begin{table}[t]
\centering
\setlength{\abovecaptionskip}{0.05cm}
\caption{Time and space overhead of range filtering index construction}
% \vspace{-7pt}
\label{tab:Range Filtered}
\resizebox{\columnwidth}{!}{
\begin{tabular}{l l c c c c c }
\toprule
\textbf{Dataset} & \textbf{Metric} & \textbf{DSG} & \textbf{IRange} & \textbf{SeRF} & \textbf{UNIFY} & \textbf{WinFilter} \\
\midrule
\multirow{3}{*}{DEEP} 
& Index time(s)  & 4156 & 1128 & \textbf{189.89} & 7510 & 32843 \\
& Index size(GB)  &  4.6 & 0.92 & \textbf{0.80} & 1.43 & 2.41  \\
& Peak memory(GB) &  7.78 & 3.54 & \textbf{1.38} & 2.61 & 7.08   \\
% \cmidrule{1-7}
\midrule
\multirow{3}{*}{YT-Audio} 
& Index time(s)   & 15314 & 1039 & \textbf{187.20} & 5905 & 30904    \\
& Index size(GB)    & 4.3 & \textbf{0.7} & 1.04 & 1.55 & 1.32  \\
& Peak memory(GB)   & 7.5 & 3.16 & \textbf{1.57} & 2.97 & 6.95  \\
\midrule
\multirow{3}{*}{WIT} 
& Index time(s)  & 29492 & 7068 & \textbf{828.12} & 28759 & 94909\\
& Index size(GB) &  4.2 & \textbf{1.5 }& 15.34 & 8.71 & 1.59  \\
& Peak memory(GB)  & 21.7 & \textbf{12.32} & 15.87 & 24.38 & 15.29\\
% \cmidrule{1-7}
\bottomrule
\end{tabular}
}
\end{table}

\textit{\textbf{Different Datasets .}}  
To evaluate the performance of various methods across different datasets, we conduct experiments on six datasets. Our analysis focuses on performance variations concerning dataset size, LID, and vector dimensionality.

@@

Similar to the evaluation of \textit{Attribute Filtering} algorithms, we also evaluate three key metrics mentioned above in \textit{Range Filtering} algorithms.

\textit{\textbf{Index construction time.}}
As shown in Table \ref{tab:Range Filtered}, SeRF achieves the fastest index construction time among the five algorithms. It incrementally builds the index in attribute order. It requires only a single pass and adds edges within local graphs. The structure is simple, with no redundant operations.

@@

\textit{\textbf{Index size.}} Across the three datasets, IRange maintains a relatively small index size. It builds independent local neighbor graphs within each interval and merges indices dynamically during queries. This prevents space explosion. On low-dimensional datasets (Deep, YT-Audio), SeRF compresses the index using sparse vector distributions, keeping it small. However, in the high-dimensional WIT dataset with high LID, vector connectivity increases. This reduces compression efficiency, leading to a sharp rise in index size. DSG maintains a stable index size across datasets but is generally larger. It uses edge structures with interval labels to support streaming insertions and range queries, increasing storage overhead.

\textit{\textbf{Peak memory usage.}}
In terms of peak memory usage, all five algorithms consume less memory on the first two datasets. However, memory usage increases significantly on WIT, likely due to its high dimensionality.

@@

In contrast, IRange achieves the lowest peak memory usage on WIT. It pre-builds graphs only for a limited number of intervals, avoiding explicit index construction for all possible query ranges. Its index structure remains stable in high dimensions without significant graph growth. This effectively controls index size, making IRange more advantageous for high-dimensional datasets.

\begin{figure*}[htbp]
  \centering
  \setlength{\abovecaptionskip}{0cm}
    \setlength{\belowcaptionskip}{-0.4cm}
  \includegraphics[width=\textwidth]{figures/exp/exp_8_2.pdf}
  \caption{RF-ANN search }
  \label{fig:exp_8_2}
\end{figure*}

\subsubsection{Performance Evaluation. }

@@

% \begin{figure*}[htbp]
%     \centering
%     \setlength{\abovecaptionskip}{0cm}
%     \setlength{\belowcaptionskip}{-0.4cm}
%     \includesvg[width=\textwidth]{figures/exp/exp_8_3.svg}
%     \caption{Dataset Effect on RF-ANN Search}
%     \label{fig:exp_8_3}
% \end{figure*}
\begin{figure*}[htbp]
    \centering
    \setlength{\abovecaptionskip}{0cm}
    \setlength{\belowcaptionskip}{-0.5cm}
    \includegraphics[width=\textwidth]{figures/exp/exp_8_3.pdf}
    \caption{Dataset Effect on RF-ANN Search}
    \label{fig:exp_8_3}
\end{figure*}

@@

\begin{table}[t]
\centering
\setlength{\abovecaptionskip}{0.05cm}
\caption{Algorithm Recommendation per Scenario}
\resizebox{\columnwidth}{!}{
\begin{tabular}{|l|c|c|}
\hline
\textbf{Scenario} & \textbf{Attribute Filtering} & \textbf{Range Filtering}\\
\hline
S1: Large-scale Datasets & Puck, UNG & iRange\\
\hline
S2: Fast Index Construction & ACORN-1, UNG  &SeRF, iRange\\
\hline
S3: Query on Hard Datasets & CAPS, UNG &iRange, DSG\\
\hline
S4: Query on Easy Datasets & StitchedVamana, NHQ & iRange\\

\hline
S5: General-purpose Use & Milvus, Faiss &DSG, UNIFY\\
\hline
S6: Resource-constrained Environments & CAPS, Faiss&SeRF, IRange \\
\hline
S7: High-Recall Querying & UNG & WinFilter \\
\hline
\end{tabular}
}
\label{tab:algo_scenarios}
\vspace{-0.1cm}
\end{table}

@@

% \end{tabular}
% \label{tab:range_algo_scenarios}
% \end{table}
Table~\ref{tab:algo_scenarios} also presents the application scenarios for range filtering. iRange demonstrates strong performance on large-scale datasets, combining efficient construction with excellent query capabilities (S1). For fast index construction, SeRF leads with iRange as a close second (S2). When handling hard datasets, iRange and DSG are recommended (S3), while iRange also performs well on easy datasets (S4). For general-purpose applications, DSG supports dynamic updates while UNIFY provides flexible filtering strategies (S5). Both SeRF and iRange show superior memory efficiency (S6). WinFilter emerges as the optimal choice for high-recall static queries (S7).

\subsection{Challenges}
\subsubsection{\textbf{Attribute Filtering}}
%修改语序
Most existing attribute filtering algorithms are graph-based, %and 
they typically outperform other methods in terms of query accuracy and efficiency. However, graph-based algorithms often incur high indexing costs and significant memory overhead. In contrast, IVF algorithms  generally consume less memory but suffer from limited performance. Filtered−DiskANN leverages disk-based storage, making it more suitable for extremely large-scale datasets even in memory-constrained environments. Moreover, the index construction phase of graph-based algorithms is computationally intensive. Therefore, exploring GPU-based acceleration for graph construction and vector computation during indexing is promising. It can significantly improve the performance of graph-based methods.
